SHARP - II: Revealing a bias in observational measurements of dark matter substructure with gravitational lens flux ratios

### 2016-01-30 papers/B1555_EdgeOn/B1555_letter.tex
B1555 advance model + paper text edit

\textbf{To obtain the best-fit lens model, we explore the parameter space as the following:  First, we use the best-fit model of \citet{Marlow99} as the starting point for the SIE profile. By looking at the NICMOS F160W imaging, which has the highest signal-to-noise ratio of disc among the data, we fix the centroid position as $(-0.162, -0.206)$ , ellipticity as $e=0.83$, and position angle of $\theta=8^\circ$. The total number of free parameters is 9. After obtaining the best-fit lens model from {\sc gravlens}, we run a Markov-Chain Monte-Carlo (MCMC) analysis to obtain uncertainties.}

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:3panel}c. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in fluxes may due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the ExpDisc component makes up about 10 per cent of the total projected mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$).  In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total enclosed mass \citep[e.g.][]{swells2}.  In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy. \textbf{[Is 0.015" still a large offset?]}
%We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.

@@

\begin{table*}
\centering
\caption{A summary of the observational constraints and the predicted lens-model parameters. The lensed image positions in the radio \citep{Marlow99} (in arcsec) and average flux ratios \citep{K03} are respect to image A. Uncertainties in the position measurements for images A, B and C are $0.001$~arcsec and for image D are $0.006$~arcsec.}
\begin{tabular}{cccccccc}
\hline
Image	&\multicolumn{3}{c}{Observed MERLIN} 	 	& \multicolumn{4}{c}{Model-Predicted}\\
		&\multicolumn{3}{c}{5 GHz}		& & & {SIE+ExpDisc} & SIE-only\\
		&East &North & Flux ratio &East 	&North & Flux ratio &Flux ratio\\ 
\hline
A  &$0$    		&$0$			&  $1$ 				&$+0.000$  &$+0.000$	& $1$ 		& $1$\\  
B  &$-0.073$	&$+0.048$	& $0.620 \pm 0.039$ 	&$-0.073$ &$+0.048$		& $0.620$ 	& $0.971$ \\  
C  &$-0.412$ 	&$-0.028$	& $0.507\pm 0.030$	&$-0.412$ &$-0.028$		& $0.461$ 	& $0.312$\\  
D  &$-0.162$	&$-0.368$	& $0.086 \pm 0.024$ 	&$-0.163$ &$-0.365$		& $0.115$ 	& $0.106$\\  
\hline
\end{tabular}
\label{tab:results}
\end{table*}

\begin{table}
\centering
\caption{Best-fit model parameters and uncertainties for our two-component SIE + exponential disc lens model (our single SIE model is effectively equivalent to the model in \citet{Marlow99} and so is not shown here). The positions are measured in arcsec, relative to the radio position of image A. The ellipticity is defined as $e=1-q$ and the position angles are measured in degrees east of north. The Einstein radius $b$ and disc scale radius are measured in arcsec.}
\begin{tabular}{ccc}
\hline 
Parameter    & SIE Component & ExpDisc Component  \\
%   & (SIE) & (ExpDisc)\\
\hline
$\Delta$RA	& $-0.176 \pm 0.005$		& $-0.162$ \textit{(fixed)}\\
$\Delta$Dec	& $-0.212 \pm 0.005$		& $-0.206$ \textit{(fixed)} \\
$b$ 			& $0.19 \pm 0.01$  		& ...  \\
$\kappa_0$ & ... &$0.19 \pm 0.01$\\
$e$	  		& $0.40 \pm 0.01$			& $0.83$ \textit{(fixed)} \\
$\theta$ 		& $4.4 \pm 0.6$			& $8.0$ \textit{(fixed)}	 \\
$R_s$			& ...  						& $0.21 \pm 0.01$	 \\
\hline
\label{tab:model}
\end{tabular}
\end{table}

@@

Here, we find evidence for an edge-on disc in the flux-ratio anomaly
gravitational lens B1555+375 that can account for the required
perturbation in the mass model.  Moreover, near-infrared AO imaging of
other lenses in the SHARP sample shows that B1555+375 is not the only
system that has a significant disc component detection in the lensing
galaxy; these gravitational lenses will be investigated in future
work, but the images alone suggest that other systems may have similar
links between flux-ratio anomalies and disc components. Such
high-order deviations from a smooth mass distribution have already
been suggested previously \citep{evans03, congdon05}, but this is the
first time that an observed disc has been explicitly modelled to
account for a flux-ratio anomaly. However, it is important to realise
that by including observationally-motivated modifications to a
`smooth' mass distribution, such as luminous dwarf companions or
edge-on discs, in the mass models, we are still limited in our
interpretation if we can not determine how much of the anomalies are
due to these components. For our current analysis of B1555+375, we
have shown that the entire anomaly could be caused by the edge-on
disc.  However, it may be possible to independently constrain the mass
of the disc through kinematic measurements (most likely from
mm-wavelength resolved emission line observations), which would test
whether the disc actually makes a sufficient contribution to the mass
model \citep[e.g. as in the case of B1933+503;][]{suyu12}. Also, in
the blue contours of Fig.~\ref{fig:3panel}a, B1555+375 shows evidence
for an extended $\sim$ 100 mas gravitational arc at radio wavelengths
that crosses the edge-on disc. Further high resolution imaging with
very long baseline interferometry will provide the observational
constraints required to fully test whether the disc is sufficient to
cause the anomaly.

Our analysis of B1555+375 demonstrates that the choice of the macromodel used to describe the `smooth' lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to a misinterpretation of the origin of flux-ratio anomalies in some systems. Before concluding that the flux-ratio anomalies in any given lens system are the result of dark matter substructure, therefore, it is critical that other possible sources of significant curvature in the lensing potential (i.e., `sharp' mass features such as edge-on disks or even massive luminous satellites) be taken into account 
\citep[e.g.][]{Xu15}. A number of lens systems with significant disc components are
already known, including, for example, the sample of 16 edge-on disks from SDSS found in the SWELLS survey \citep{swells1}.  This result is timely, as upcoming large-scale surveys (e.g., the Dark Energy Survey, Kilo-Degree Survey, Euclid, and the Large Synoptic Survey Telescope) will deliver orders of magnitude more gravitationally lensed quasars for which flux ratios can
  be measured using narrow lines \citep[e.g.][]{N14}.
  %Finally, we note that massive edge-on discs are \textit{not} observed in some flux-ratio anomaly systems \citep[and the presence of a disc-like component may not resolve the anomaly anyway;][]{evans03}, indicating that these systems are still useful for probing the dark matter substructure mass function.

@@

\bibliographystyle{mnras}
\bibliography{reference.bib}

>>>>

\textbf{We use {\sc gravlens} to obtain the best-fit SIE+ExpDisc lens model. The best-fit parameters from \citet{Marlow99} are set as the starting point of the SIE profile. By looking at the NICMOS F160W imaging, which has the highest signal-to-noise ratio of disc among the data, we fix the centroid position as $(-0.162, -0.206)$ , ellipticity as $e=0.83$, and position angle of $\theta=8^\circ$. The total number of free parameters is 9. After obtaining the best-fit lens model, we run a Markov-Chain Monte-Carlo (MCMC) analysis to obtain uncertainties.}

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:3panel}c. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in fluxes may due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the \textbf{ExpDisc component makes up about 15 per cent of the total projected mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$).  In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total enclosed mass \citep[e.g.][]{swells2, swells5}.}  In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy. 
%We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.

@@

\begin{table*}
\centering
\caption{A summary of the observational constraints and the predicted lens-model parameters. The lensed image positions in the radio \citep{Marlow99} (in arcsec) and average flux ratios \citep{K03} are respect to image A. Uncertainties in the position measurements for images A, B and C are $0.001$~arcsec and for image D are $0.006$~arcsec.}
\begin{tabular}{cccccccc}
\hline
Image	&\multicolumn{3}{c}{Observed MERLIN} 	 	& \multicolumn{4}{c}{Model-Predicted}\\
		&\multicolumn{3}{c}{5 GHz}		& & & {SIE+ExpDisc} & SIE-only\\
		&East &North & Flux ratio &East 	&North & Flux ratio &Flux ratio\\ 
\hline
A  &$0$    		&$0$			&  $1$ 				&$+0.000$  &$+0.000$	& $1$ 		& $1$\\  
B  &$-0.0726$	&$+0.0480$	& $0.620 \pm 0.039$ 	&$-0.0715$ &$+0.0487$		& $0.606$ 	& $0.971$ \\  
C  &$-0.4117$ 	&$-0.0280$	& $0.507\pm 0.030$	&$-0.4115$ &$-0.0278$		& $0.445$ 	& $0.312$\\  
D  &$-0.1619$	&$-0.3680$	& $0.086 \pm 0.024$ 	&$-0.1693$ &$-0.3704$		& $0.123$ 	& $0.106$\\  
\hline
\end{tabular}
\label{tab:results}
\end{table*}

\begin{table}
\centering
\caption{Best-fit model parameters and uncertainties for our two-component SIE + exponential disc lens model (our single SIE model is effectively equivalent to the model in \citet{Marlow99} and so is not shown here). The positions are measured in arcsec, relative to the radio position of image A. The ellipticity is defined as $e=1-q$ and the position angles are measured in degrees east of north. The Einstein radius $b$ and disc scale radius are measured in arcsec.}
\begin{tabular}{ccc}
\hline 
Parameter    & SIE Component & ExpDisc Component  \\
%   & (SIE) & (ExpDisc)\\
\hline
$\Delta$RA	& $-0.177 \pm 0.005$		& $-0.162$ \textit{(fixed)}\\
$\Delta$Dec	& $-0.205 \pm 0.005$		& $-0.206$ \textit{(fixed)} \\
$b$ 			& $0.18 \pm 0.01$  		& ...  \\
$\kappa_0$ & ... &$0.27 \pm 0.01$\\
$e$	  		& $0.25 \pm 0.01$			& $0.83$ \textit{(fixed)} \\
$\theta$ 		& $4.9 \pm 0.6$			& $8.0$ \textit{(fixed)}	 \\
$R_s$			& ...  						& $0.24 \pm 0.01$	 \\
\hline
\label{tab:model}
\end{tabular}
\end{table}

@@

Here, we find evidence for an edge-on disc in the flux-ratio anomaly
gravitational lens B1555+375 that can account for the required
perturbation in the mass model.  \textbf{Although the constraint on disc mass from photometry is currently not avaliable due to the low signal-to-noise ratio in AO imaging and the lack of redshift information, our model indicates that the flux anomalies in B1555+375, at least, have strong link to the present of its edge-on disc. } Such
high-order deviations from a smooth mass distribution have already
been suggested previously \citep{evans03, congdon05}, but this is the
first time that an observed disc has been explicitly modelled to
account for a flux-ratio anomaly. However, it is important to realise
that by including observationally-motivated modifications to a
`smooth' mass distribution, such as luminous dwarf companions or
edge-on discs, in the mass models, we are still limited in our
interpretation if we can not determine how much of the anomalies are
due to these components. For our current analysis of B1555+375, we
have shown that the entire anomaly could be caused by the edge-on
disc.  However, it may be possible to independently constrain the mass
of the disc through kinematic measurements (most likely from
mm-wavelength resolved emission line observations), which would test
whether the disc actually makes a sufficient contribution to the mass
model \citep[e.g. as in the case of B1933+503;][]{suyu12}. Also, in
the blue contours of Fig.~\ref{fig:3panel}a, B1555+375 shows evidence
for an extended $\sim$ 100 mas gravitational arc at radio wavelengths
that crosses the edge-on disc. Further high resolution imaging with
very long baseline interferometry will provide the observational
constraints required to fully test whether the disc is sufficient to
cause the anomaly.

Our analysis of B1555+375 demonstrates that the choice of the macromodel used to describe the `smooth' lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to a misinterpretation of the origin of flux-ratio anomalies in some systems. \textbf{\citet{Maller97,Kee98, Moller03} have shown that a disc with high inclination angle can significantly affects the lensing cross-section. While the current sample of lensed quasars with flux ratio anomalies is fairly small, few systems with edge-on disc is sufficient to bring a bias into the substructure analysis. Moreover, near-infrared AO imaging of
other lenses in the SHARP sample shows that B1555+375 is not the only
system that has a significant disc component detection close to merging images among the current sample of flux ratio anomalous systems; these gravitational lenses will be investigated in future
work, but the images alone suggest that other systems may have similar
links between flux-ratio anomalies and disc components. This bias brought by edge-on disc may be avoided by post-selection as the sample size grows in the future. Therefore, it is critical to obtain high-resolution imaging like SHARP of target lenses to detect `sharp' mass features such as edge-on disks or even massive luminous satellites.} 
% A number of lens systems with significant disc components arealready known, including, for example, the sample of 16 edge-on disks from SDSS found in the SWELLS survey \citep{swells1}. 
 This result is timely, as upcoming large-scale surveys (e.g., the Dark Energy Survey, Kilo-Degree Survey, Euclid, and the Large Synoptic Survey Telescope) will deliver orders of magnitude more gravitationally lensed quasars for which flux ratios can
  be measured using narrow lines \citep[e.g.][]{N14}.
  %Finally, we note that massive edge-on discs are \textit{not} observed in some flux-ratio anomaly systems \citep[and the presence of a disc-like component may not resolve the anomaly anyway;][]{evans03}, indicating that these systems are still useful for probing the dark matter substructure mass function.

@@

\bibliographystyle{mnras}
\bibliography{reference}
